---
title: "Exploratory Data Analysis"
author: "STOR 320 Group 4"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# Import Data Below
# read_csv()
```

#Creator: Katharine Kurz

###Q1: What regions of the country have the highest rates of police killings?

```{r}
#
```

###Q2: Which races/ethnicities are frequently subjected to police violence?

```{r}
#
```

#Interpreter: Chris Jakuc

###Q1: Which age groups are more likely to be affected by police killings?

```{r}
#
```

###Q2: What role does income a factor in the frequency of police killings by region/state?


```{r}
#
```

#Orator: Harrison Cho

###Q1: What factors combined are most influential in predicting police killings in a city?

```{r}
#
```

###Q2: Are law enforcement agencies that hold the highest killing rates primarily influenced by population size or other variables? 

```{r}
#
```

#Orator: Mohammad Aamin

###Q1: On a national level, does a relationship exist between the proportion of those who graduated college and the number of police killings?

```{r}
#
```

###Q2: Which uncommon police weapons result in fatalities at a higher frequency?

```{r}
#
```

#Deliverer: Avinash Gandhi

###Q1: Do areas with higher unemployment rates tend to have higher rates of police killings?

```{r}
#
```

###Q2: Were more armed victims killed than unarmed victims?

```{r}
#
```

# Follow up questions

### New Questions Based Off Initial Investigation

- Q1: WRITE_QUESTION_HERE
- Q2: WRITE_QUESTION_HERE
- Q3: WRITE_QUESTION_HERE 
- Q4: WRITE_QUESTION_HERE

### Investigation of Follow-up Questions

GIVE WHAT 2 QUESTIONS YOU ATTEMPTED TO INVESTIGATE FURTHER IN COMPLETE SENTENCES (*Example:* Our group decided to investigate Q2 and Q4 in further detail.)

SHOW AT LEAST 2 TABLES OR FIGURES BELOW THAT EXPLORE ANSWERS FOR THE QUESTIONS YOU ARE INVESTIGATING FURTHER.

```{r}
#
```

```{r}
#
```

#Summary

GIVE A 2 PARAGRAPH SUMMARY. 

PARAGRAPH 1 SHOULD DESCRIBE WHAT YOU LEARNED ABOUT YOUR DATA FROM INVESTIGATING THE INITIAL QUESTIONS. DID YOU FIND ANYTHING UNUSUAL IN YOUR DATA? DID ANYTHING SURPRISE YOU? WHICH OF THE INITIAL QUESTIONS WERE HELPFUL IN LEADING YOU TO MORE QUESTIONS?

PARAGRAPH 2 SHOULD SUMMARIZE WHAT YOU LEARNED FROM INVESTIGATING THE FOLLOW-UP QUESTIONS. WHY ARE THESE FOLLOW-UP QUESTIONS INTERESTING FOR INVESTIGATION? DESCRIBE THE TABLES/FIGURES YOU USED TO EXPLORE ANSWERS TO THESE FOLLOW-UP QUESTIONS? WHAT DID YOU LEARN FROM THE TABLES/FIGURES REGARDING THE FOLLOW-UP QUESTIONS YOU PROPOSED?















