---
title: "Exploratory Data Analysis"
author: "STOR 320 Group 4"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(Hmisc)
library(modelr)
library(dplyr)
library(xtable)
library(forcats)
library(formattable)
library(rvest)


# Import Data Below
PoliceKillings<-read_csv("PKcleanP2.csv")

# Add DC to the South region
PoliceKillings$region[which(PoliceKillings$State=="DC")]<-"South"
```

#Creator: Katharine Kurz


###Q1: Which regions of the country have the highest rates of police killings?

```{r}
# Adding populations by state and adding populations by region
populations<-read_csv("StatePops.csv")
pops<-populations[-1,3:11]


# Import states by region
region.data<-read_csv("StatesbyCensusRegion.csv")


# Joining state pop and region data
pops1<-pops %>%
  select("GEO.display-label","respop72015") %>%
  rename(State="GEO.display-label",Pop2015=respop72015) %>%
  left_join(region.data,by="State") %>%
  mutate(Pop2015=as.numeric(Pop2015)) %>%
  group_by(region) %>%
  mutate(PopByRegion=sum(Pop2015)) %>%
  select(PopByRegion,region) %>%
  ungroup() %>%
  mutate(Pop.prop=PopByRegion/PopByRegion[1])

# Create a tibble of uniqe regions, pops, and prop pops
pops2<- tibble(PopByRegion=unique(pops1$PopByRegion),Region=unique(pops1$region),Pop.prop=unique(pops1$Pop.prop))

# Converting region variable to a factor
regions<-pops2 %>%
  mutate(Region=as_factor(Region))
levels(regions$Region)

# Remove the row for US total police killings
regions<-regions[-c(is.na(regions$Region)),]

```

```{r}
# Print out a plot of proportion of population in each region
ggplot(regions) +
  geom_col(aes(x=fct_explicit_na(Region,na_level = "US Total"),y=Pop.prop))+
  xlab("Region")+
  ylab("Proportion of Population")+
  ggtitle("Proportion of Total Population by Region")+
  theme_minimal()
```

```{r}
# Count how many police killings are in each state
state.kills<-PoliceKillings %>%
  select(State) %>%
  mutate(State=as_factor(State)) %>%
  group_by(State) %>%
  dplyr::summarize(count=n())

# Count how many police killings are in each region
region.kills<-PoliceKillings %>%
  select(region) %>%
  mutate(region=as_factor(region)) %>%
  group_by(region) %>%
  dplyr::summarize(count=n())

```

```{r}
# Create population state data frame
states<- pops %>%
  select("GEO.display-label","respop72015") %>%
  rename(State="GEO.display-label",Pop2015=respop72015) %>%
  mutate(Pop2015=as.numeric(Pop2015)) %>%
  left_join(region.data,by="State") %>%
  select(-region)

# Join police killings by state with population state data and create a variable for kills per 100,000 per state
states1<-states[-c(1),] %>%
  select(Pop2015,State,state) %>%
  right_join(state.kills,by=c("state"="State")) %>%
  mutate(kill.pc.state=count/Pop2015*100000)

# Join police killings by region with population regions data and create a variable for kills per 100,000 per region
regions1<-regions %>%
  select(PopByRegion,Region) %>%
  right_join(region.kills,by=c("Region"="region")) %>%
  mutate(kill.pc.region=count/PopByRegion*100000)

```

```{r}
# Create a graph for police killing rates per 100,000 people by region
ggplot(regions1)+
  geom_col(aes(x=Region,y=kill.pc.region))+
  ylab("Kills Per 100,000 People")+
  ggtitle("Kills per 100,000 People by Region")
```

###Q2: Which races/ethnicities are frequently subjected to police violence?

```{r}
Races <- PoliceKillings %>% 
  group_by(Race,region) %>%
  dplyr::summarize(count=n()) %>%
  ungroup() %>%
  mutate(prop=count/sum(count)) 

# Remove the row for unknown races
Races<-Races[-c(which(Races$Race=="Unknown")),]

# Plotting proportion of total police killings by race and religion
ggplot(Races) +
  geom_col(aes(x=Race,y=prop))+
  facet_wrap(region~.)+
  ylab("Proportion of total police killings")+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  ggtitle("Proportion of Total Police Killings by Race and Region")

# Plotting proportion of police killings in each region by race
Races1<-Races %>%
  left_join(region.kills,by="region") %>%
  mutate(Prop.race.region=count.x/count.y)



ggplot(Races1)+
  geom_col(aes(x=Race,y=Prop.race.region))+
  facet_wrap(region~.)+
  ylab("Proportion of police killings by region")+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  ggtitle("Proportion of Police Killings in Each Region by Race")

```

#Interpreter: Chris Jakuc

###Q1: Which age groups are more likely to be affected by police killings?

```{r}
Ages <- PoliceKillings %>% 
  mutate(Age = as.factor(Age)) %>%
  group_by(Age)
levels(Ages$Age)
```

```{r}
Ages1 <- Ages %>%
  mutate(Age_Bin=fct_collapse(Age,
                           Teenager = levels(Age)[c(1:4)],
                           Twenties = levels(Age)[c(5:14)],
                           Thirties = levels(Age)[c(15:25)],
                           Forties = levels(Age)[c(26:34)],
                           Fifties = levels(Age)[c(35:44)],
                           Sixties = levels(Age)[c(45:52)],
                           Seventies =  levels(Age)[c(53:58)],
                           Eighties = levels(Age)[c(59:60)],
                           Unknown = levels(Age)[c(61)]
                           )) %>%
           group_by(Age_Bin) %>%
           dplyr::summarize(n=n()) %>%
           ungroup() %>%
           mutate(prop=n/sum(n))
           
# Remove the row for unknown age
Ages1<-Ages1[-c(which(Ages1$Age_Bin=="Unknown")),]
```
group ranges -
16-19 - Teenager
20-29 - Twenties
30-39 - Thirties
40-49 - Forties
50 -59 - Fifties
60 - 69 - Sixties
70-79 - Seventies
80-89 - Eighties
Unknown - 

```{r}
ggplot(Ages1)+
  geom_col(aes(x=Age_Bin,y=prop))+
  xlab("Age Group")+
  ylab("Proportion of Total Killings")+
  ggtitle("Proportion of Total Killings by Age Group")+
  theme_minimal()
```

###Q2: What role does income play in the frequency of police killings by region/state?


```{r}
mean.inc<-mean(PoliceKillings$med.tract.household.inc)

ggplot(PoliceKillings)+
  geom_histogram(aes(x=med.tract.household.inc),bins=20,fill='deepskyblue1')+
  geom_vline(xintercept=mean.inc,linetype="dashed",color="red")+
  facet_wrap(region~.)+
  theme_dark()+
  xlab("Median Household Income")+
  ylab("Number of Deaths")+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  scale_x_continuous(breaks=c(25000,50000,75000,100000))
```

#Orator: Harrison Cho

###Q1: What factors combined are most influential in predicting police killings in a city?

```{r}
PD = 
  PoliceKillings %>% 
  mutate( 
    killed = ifelse("age">16, 1, 0), 
    killed = factor(killed, levels=c("0","1"), labels = c("Alive","Dead")))
  
ggplot(PD) + 
  geom_bar(mapping = aes(killed)) + 
  theme_light() + 
  ggtitle("Police Violence in 2015") + xlab("Status of Victims") + ylab("Count")+
  scale_x_discrete(limits=c("Alive","Dead"))
```

This question is unable to be answered with the current dataset. Given that everyone is killed, we are unable to answer this question using a model. 

###Q2: Are law enforcement agencies that hold the highest killing rates primarily influenced by population size or other variables? 

```{r}
Agencies = 
  PoliceKillings %>% 
  arrange(Law_Enforcement_Agency) %>% 
  group_by(Law_Enforcement_Agency) %>%
  dplyr::summarize(n = n()) %>%
  mutate( freq = n / sum(n))

Agencies2 = inner_join(Agencies, PoliceKillings) %>%
  select("Law_Enforcement_Agency","freq", -("n"), -("Name"), "Age":"region") 

```

```{r}
Agencies = 
  PoliceKillings %>% 
  group_by(Law_Enforcement_Agency) %>%
  dplyr::summarize(n = n()) %>%
  mutate( freq = n / sum(n))

Agencies2 = inner_join(Agencies, PoliceKillings) %>%
  select("Law_Enforcement_Agency","freq", -("n"), -("Name"), "Age":"region") 

Agencies3 = 
  Agencies2 %>% 
  group_by(Law_Enforcement_Agency) %>% 
  mutate(Population = median(Population),
         Proportion_White = median(Proportion_White), 
         Proportion_Black = median(Proportion_Black), 
         Proportion_Hispanic = median(Proportion_Hispanic), 
         med.tract.personal.inc = median(med.tract.personal.inc),
         med.tract.household.inc = median(med.tract.household.inc), 
         med.county.household.inc = median(med.county.household.inc), 
         Poverty_Rate = median(Poverty_Rate),
         Unemployment_Rate = median(Unemployment_Rate),
         Proportion_Attended_College = median(Proportion_Attended_College), 
         inc.quint.nat = median(inc.quint.nat),
         tractvcount.percentile = median(tractvcount.percentile),
         Age = as.numeric(Age),
         Age = median(Age))

ggplot(data = Agencies3, mapping = aes(x = Population, y=freq, color = region)) + geom_point() + ylab("Median Population") + ylab("Frequency of Fatality") + ggtitle("Frequencies of Fatal Police Shootings Grouped by Department")
```


#Orator: Mohammad Aamin

###Q1: On a national level, does a relationship exist between the proportion of those who graduated college and the number of police killings?

```{r}
ggplot(data=PoliceKillings) + 
  geom_histogram(mapping=aes(x=Proportion_Attended_College),bins=15,fill="deepskyblue1")+
  labs(x="Proportion that Attended College",y="Number of Deaths",title="College Education vs. Police Killings") + theme_dark()
```

###Q2: Which uncommon police weapons result in fatalities at a higher frequency?

```{r}
unique(PoliceKillings$Cause_Of_Death)
ggplot(data=PoliceKillings) +
  geom_bar(aes(x=Cause_Of_Death))
NoGun= filter(PoliceKillings, Cause_Of_Death != "Gunshot")
ggplot(data=NoGun) +
  geom_bar(aes(x=Cause_Of_Death), color="black",fill=c("red","blue","darkgreen","purple"),size=2) + 
  xlab("Cause of Death") + ylab("Number of Deaths") + 
  theme_light()
```

#Deliverer: Avinash Gandhi

###Q1: Do areas with higher unemployment rates tend to have higher rates of police killings?

From the dataset it can be observed that most of the police killings occured in areas that had an unemployment above the national rate in  the second quarter of 2015 which was 5.4%. 400 of the 465 killings occured in areas that had unemployment rates above the national average at that time meaning that higher unemployment rates could have a significant impact on where police killings occur. Additionally 235 of the killings occured in areas that had an unemployment rate greater than 10%, further substantiating the theory that high unemployment and police killings are correlated.
```{r}

#count of whether each killing occured in an area that has an unemployment rate higher than the national average of 3.6%
PK1=PoliceKillings%>%
  mutate(higherthanavg=if_else(Unemployment_Rate>0.054,"Yes","No"))
table(PK1$higherthanavg) 

#splitting up unemployment into ranges and then finding the count for each bin
URranges1=cut2(PoliceKillings$Unemployment_Rate,c(0.02,0.04,0.06,0.08,0.1,0.12,0.14,0.16,0.18,0.2,0.22,0.24,0.26,0.28,0.3,0.32,0.34,0.36,0.38,0.4)) 
summary(URranges1) 


#histogram of unemployment values
ggplot(PoliceKillings,aes(x=Unemployment_Rate))+geom_histogram(breaks=c(0.00,
                      seq(0.05,0.5,0.05)),col="magenta")+geom_vline(xintercept = 0.054,linetype="dashed",color="yellow")



```

###Q2: Were more armed victims killed than unarmed victims?
From the data it can be observed that 357 of the 465 people killed by the police in the first half of 2015 were armed, while 102 were unarmed, and for 7 of the individuals it is unknown whether they were armed. From this breakdown it seems likely that being armed is a major factor in whether someone is killed by the police as there were more than three times the number of armed individuals as unarmed in the dataset.

```{r}
#Creating a table of killings by category of count

Armed1=table(PoliceKillings$Armed)
print(Armed1)
ggplot(PoliceKillings,aes(x=PoliceKillings$Armed))+geom_bar()+coord_flip()+xlab("Armed vs Unarmed")
```

# Follow up questions

### New Questions Based Off Initial Investigation

- Q1: Which model best predicts the median household income of a police killing victim?
- Q2: How do gun laws relate to observations of police killings?
- Q3: How do frequencies of police killings by cause of death proportionally relate to region/gun law groupings? 
- Q4: How do frequencies of police killings by armed status proportionally relate to region/gun law groupings?

### Investigation of Follow-up Questions

GIVE WHAT 2 QUESTIONS YOU ATTEMPTED TO INVESTIGATE FURTHER IN COMPLETE SENTENCES (*Example:* Our group decided to investigate Q2 and Q4 in further detail.)

SHOW AT LEAST 2 TABLES OR FIGURES BELOW THAT EXPLORE ANSWERS FOR THE QUESTIONS YOU ARE INVESTIGATING FURTHER.

## Question 1: Which model best predicts the median household income of a police killing victim?

```{r}
library(tidyverse)    #Essential Functions
library(modelr)       #Helpful Functions in Modeling
library(purrr)
library(broom)
```

```{r}
income_cv1 = PoliceKillings %>% 
  select(-Name, -Latitude, -Longitude,everything()) %>% 
  mutate( 
    Age = as.numeric(Age),
    Gender = as.factor(Gender), 
    Race = as.factor(Race), 
    State = as.factor(State),
    Region = as.factor(region)) 
```

```{r}
firstmod = lm(med.tract.household.inc~Population,data=income_cv1)
tidy(firstmod)

income_cv1_sec = income_cv1 %>% 
  crossv_kfold(5)

head(income_cv1_sec)

train.model.func1 = function(data){
  fistmod = lm(med.tract.household.inc~Population,data=data)
  return(fistmod)
}
income_cv1_third = income_cv1_sec %>%
  mutate(tr.model = map(train,train.model.func1)) 

income_cv1_predict = income_cv1_third %>% 
  mutate(predict = map2(test,tr.model, ~augment(.y,newdata=.x))) %>% 
  select(predict) %>% 
  unnest()

head(income_cv1_predict)

bias.func=function(actual,predict){
  bias=mean((actual-predict),na.rm=T)
  return(bias)
}

mae.func=function(actual,predict){
  mae=mean(abs(actual-predict),na.rm=T)
  return(mae)
}

RMSE.func=function(actual,predict){
  mse=mean((actual-predict)^2,na.rm=T)
  rmse=sqrt(mse)
  return(rmse)
}
```


```{r}
secmod = lm(med.tract.household.inc~Population + Poverty_Rate, data=income_cv1)
tidy(secmod)

income_cv2_sec = income_cv1 %>% 
  crossv_kfold(5)

train.model.func1 = function(data){
  secmod = lm(med.tract.household.inc~Population + Poverty_Rate, data=income_cv1)
  return(secmod)
}
income_cv2_third = income_cv2_sec %>%
  mutate(tr.model = map(train,train.model.func1)) 

income_cv2_predict = income_cv2_third %>% 
  mutate(predict = map2(test,tr.model, ~augment(.y,newdata=.x))) %>% 
  select(predict) %>% 
  unnest()
```

```{r}
fourmod = lm(med.tract.household.inc~Population + Poverty_Rate + Age, data=income_cv1)
tidy(fourmod)
summary(fourmod)

income_cv4_sec = income_cv1 %>% 
  crossv_kfold(100)

train.model.func1 = function(data){
  fourmod = lm(med.tract.household.inc~Population + Poverty_Rate + Age, data=income_cv1)
  return(fourmod)
}
income_cv4_third = income_cv4_sec %>%
  mutate(tr.model = map(train,train.model.func1)) 

income_cv4_predict = income_cv4_third %>% 
  mutate(predict = map2(test,tr.model, ~augment(.y,newdata=.x))) %>% 
  select(predict) %>% 
  unnest()
```

```{r}
fivemod = lm(med.tract.household.inc~Population + Poverty_Rate + Age + Proportion_Attended_College, data=income_cv1)
tidy(fivemod)
summary(fivemod)

income_cv5_sec = income_cv1 %>% 
  crossv_kfold(100)

train.model.func1 = function(data){
  fivemod = lm(med.tract.household.inc~Population + Poverty_Rate + Age + Proportion_Attended_College, data=income_cv1)
  return(fivemod)
}
income_cv5_third = income_cv5_sec %>%
  mutate(tr.model = map(train,train.model.func1)) 

income_cv5_predict = income_cv5_third %>% 
  mutate(predict = map2(test,tr.model, ~augment(.y,newdata=.x))) %>% 
  select(predict) %>% 
  unnest()
```

```{r}
sixmod = lm(med.tract.household.inc~Population + Poverty_Rate + Age + Proportion_Attended_College + Proportion_White + Race + Region + Race*Region, data=income_cv1)
tidy(sixmod)
summary(sixmod)

income_cv6_sec = income_cv1 %>% 
  crossv_kfold(100)

train.model.func1 = function(data){
  sixmod = lm(med.tract.household.inc~Population + Poverty_Rate + Age + Proportion_Attended_College + Proportion_White + Race + Region + Race*Region, data=income_cv1)
tidy(sixmod)
  return(sixmod)
}
income_cv6_third = income_cv6_sec %>%
  mutate(tr.model = map(train,train.model.func1)) 

income_cv6_predict = income_cv6_third %>% 
  mutate(predict = map2(test,tr.model, ~augment(.y,newdata=.x))) %>% 
  select(predict) %>% 
  unnest()
```

```{r, results="asis"}
table <- matrix(c(bias.func(actual=income_cv1_predict$med.tract.household.inc, predict=income_cv1_predict$.fitted), mae.func(actual=income_cv1_predict$med.tract.household.inc, predict=income_cv1_predict$.fitted), RMSE.func(actual=income_cv1_predict$med.tract.household.inc, predict=income_cv1_predict$.fitted), bias.func(actual=income_cv2_predict$med.tract.household.inc, predict=income_cv2_predict$.fitted), mae.func(actual=income_cv2_predict$med.tract.household.inc, predict=income_cv2_predict$.fitted), RMSE.func(actual=income_cv2_predict$med.tract.household.inc, predict=income_cv2_predict$.fitted), bias.func(actual=income_cv4_predict$med.tract.household.inc, predict=income_cv4_predict$.fitted), mae.func(actual=income_cv4_predict$med.tract.household.inc, predict=income_cv4_predict$.fitted), RMSE.func(actual=income_cv4_predict$med.tract.household.inc, predict=income_cv4_predict$.fitted),bias.func(actual=income_cv5_predict$med.tract.household.inc, predict=income_cv5_predict$.fitted), mae.func(actual=income_cv5_predict$med.tract.household.inc, predict=income_cv5_predict$.fitted), RMSE.func(actual=income_cv5_predict$med.tract.household.inc, predict=income_cv5_predict$.fitted), bias.func(actual=income_cv6_predict$med.tract.household.inc, predict=income_cv6_predict$.fitted), mae.func(actual=income_cv6_predict$med.tract.household.inc, predict=income_cv6_predict$.fitted),RMSE.func(actual=income_cv6_predict$med.tract.household.inc, predict=income_cv6_predict$.fitted)), ncol = 3, byrow = T)

colnames(table) = c("Bias", "MAE", "RMSE")
row.names(table) = c("Base Model", "Model 2", "Model 3", "Model 4", "Model 5")
table <- as.table(table) 
```

```{r}
table2=xtable(table,digits=4,align=c("l","l","l","l"))
formattable(table2)
```

## Question 2: How do gun laws relate to observations of police killings?

```{r}

# Web scrape the gun laws table from
URL.gun.scorecard = "https://lawcenter.giffords.org/scorecard2015/"
gun.scorecard = URL.gun.scorecard %>%
                      read_html() %>%
                      html_table(fill=1) %>%
                      .[[1]]

# Removing the minus and plus signs from the grades
gun.scorecard$`2015 Grade`[gun.scorecard$`2015 Grade` %in% "A-"]="A"
gun.scorecard$`2015 Grade`[gun.scorecard$`2015 Grade` %in% "B+"]="B"
gun.scorecard$`2015 Grade`[gun.scorecard$`2015 Grade` %in% "B-"]="B"
gun.scorecard$`2015 Grade`[gun.scorecard$`2015 Grade` %in% "C-"]="C"
gun.scorecard$`2015 Grade`[gun.scorecard$`2015 Grade` %in% "D-"]="D"

# Adding the region
gun.scorecard<-gun.scorecard %>%
  left_join(region.data,by="State")
 

gun.scorecard<-pops %>%
  select("GEO.display-label","respop72015") %>%
  rename(State="GEO.display-label",Pop2015=respop72015) %>%
  left_join(gun.scorecard,by="State") %>% 
  mutate(Pop2015=as.numeric(Pop2015)) %>%
  left_join(state.kills,by=c("state"="State"))

# Remove DC and Total US
gun.scorecard1<-gun.scorecard[-c(which(is.na(gun.scorecard$`2015 Grade`))),] %>%
  select(-`Death Rate Per 100k`,-`New Smart Gun Laws in 2015`)

# Turning the gun law grades into factors to run a regression later
gun.scorecard2<-gun.scorecard1[-c(which(is.na(gun.scorecard1$count))),]

gun.scorecard3<-gun.scorecard2 %>%
  mutate("2015 Grade"=factor(gun.scorecard2$`2015 Grade`)) %>%
  mutate(KillCount=count) %>%
  select(-count)

# Find how many unique groups of gun laws there are and how many kills per capita are in each
gun.scorecard4<-gun.scorecard3 %>%
  group_by(`2015 Grade`) %>%
  dplyr::summarize(count.gunlaws.ranks=n(),
                   KillCount=sum(KillCount),
                   Population=sum(Pop2015),
                   "Kill.per.100,000"=KillCount/Population*100000) %>%
  ungroup()
 
ggplot(gun.scorecard4)+
  geom_col(aes(x=`2015 Grade`,y=`Kill.per.100,000`),fill="red")+
  xlab("Gun Control Law Rankings")+
  ylab("Police Killings per 100,000 People")+
  ggtitle("Police Killings per 100,000 People by Type of Gun Laws")+
  theme_dark()+
  theme_minimal()


formattable(gun.scorecard4)

```

```{r}
library(mapproj)
library(usmap)
library(ggplot2)
library(mapdata)
library(sf)

PoliceKillingsMap = 
  PoliceKillings %>% 
  filter(State != "HI") %>%
  filter(State != "AK")

labs1 <- data.frame(
long = PoliceKillingsMap$Longitude, 
lat = PoliceKillingsMap$Latitude, 
state = PoliceKillingsMap$State,
stringsAsFactors = FALSE)


us_states1 = map_data("state")

us_states2 = 
  us_states1 %>% 
  mutate(State = region) %>% 
  select(-subregion,-region)

gun.scorecard2[[1]] = tolower(gun.scorecard2[[1]])

us_states3 = right_join(us_states2,gun.scorecard2)
```

```{r include = F}
rename(us_states3, Rank = "2015 Grade") 
```

```{r}
labs2 = right_join(us_states3, labs1)


one = ggplot(data = us_states3, aes(x =long, y = lat, group = group, fill = Rank)) + guides(fill = FALSE) + coord_fixed(1.3)

two = one + geom_polygon(color = "gray90", size = 0.1) + 
  coord_map(projection = "albers", lat0 = 39, lat1 = 45)

three = two + scale_fill_gradient(low = "white", high = "#CB454A")


four = three +
  geom_point(data = labs2, aes(x = long, y = lat))

four
```

## Question 3: How do frequencies of police killings by cause of death proportionally relate to region/gun law groupings?

```{r}



gun.scorecard.causes<-pops %>%
  select("GEO.display-label") %>%
  rename(State="GEO.display-label") %>%
  left_join(gun.scorecard,by="State") %>% 
  mutate(Pop2015=as.numeric(Pop2015)) %>%
  left_join(state.kills,by=c("state"="State"))

  # Remove DC and Total US
gun.scorecard2<-gun.scorecard1[-c(which(is.na(gun.scorecard1$`2015 Grade`))),]


# Turning the gun law grades into factors to run a regression later
gun.scorecard3<-gun.scorecard2[-c(which(is.na(gun.scorecard2$count))),]

gun.scorecard4<-gun.scorecard3 %>%
  mutate("2015 Grade"=factor(gun.scorecard3$`2015 Grade`)) %>%
  mutate(KillCount=count) %>%
  select(-count)

# Find how many unique groups of gun laws there are and how many kills per capita are in each
gun.scorecard5<-gun.scorecard4 %>%
  group_by(`2015 Grade`) %>%
  dplyr::summarize(count.gunlaws.ranks=n(),
                   KillCount=sum(KillCount),
                   Population=sum(Pop2015),
                   "Kill.per.100,000"=KillCount/Population*100000) %>%
  ungroup()



# Plotting proportion of causes of police killings in each region
Causes <- PoliceKillings %>% 
  group_by(Cause_Of_Death,region) %>%
  dplyr::summarize(count=n()) %>%
  ungroup() %>%
  mutate(prop=count/sum(count)) 

Causes1<-Causes %>%
  left_join(region.kills,by="region") %>%
  mutate(Prop.cause.region=count.x/count.y)

NoGun1= filter(Causes1, Cause_Of_Death != "Gunshot", Cause_Of_Death != "Unknown")
ggplot(data=NoGun1) +
  geom_col(aes(x=Cause_Of_Death,y=Prop.cause.region)) + 
  xlab("Cause of Death") + ylab("Proportional of Deaths") + 
  theme_light()+
  facet_wrap(NoGun1$region~.)+
  theme(axis.text.x=element_text(angle=45,hjust=1))

## Adding gun law ranking

# Count police killings by gun law rankings
laws.kills<-PoliceKillings %>%
  left_join(gun.scorecard,by="region") %>%
  select(`2015 Grade`,Cause_Of_Death) %>%
  mutate(`2015 Grade`=factor(`2015 Grade`)) %>%
  group_by(`2015 Grade`) %>%
  dplyr::summarize(Kills.By.Laws=n()) %>%
  ungroup() 
  

causes.laws<-PoliceKillings %>%
  select(Cause_Of_Death,region) %>%
  left_join(gun.scorecard,by="region") %>%
  select(Cause_Of_Death,`2015 Grade`) %>%
  group_by(Cause_Of_Death,`2015 Grade`) %>%
  dplyr::summarize(count=n()) %>%
  ungroup() %>%
  left_join(laws.kills,by="2015 Grade") %>%
  mutate(cause.prop.by.laws=count/Kills.By.Laws) %>%
  group_by(Cause_Of_Death,`2015 Grade`)

# Figure of proportional deaths by cause and gun laws
ggplot(data=causes.laws) +
  geom_col(aes(x=Cause_Of_Death,y=cause.prop.by.laws)) + 
  xlab("Cause of Death") + ylab("Proportion of Deaths") + 
  theme_light()+
  facet_wrap(causes.laws$`2015 Grade`~.)+
  theme(axis.text.x=element_text(angle=45,hjust=1))


```

## Question 4: How do frequencies of police killings by armed status proportionally relate to region/gun law groupings?


```{r}
# Plotting proportion of people armed/unarmed in police killings in each region
Armed <- PoliceKillings %>% 
  group_by(Armed,region) %>%
  dplyr::summarize(count=n()) %>%
  ungroup() %>%
  mutate(prop=count/sum(count)) 

Armed1<-Armed %>%
  left_join(region.kills,by="region") %>%
  mutate(Prop.armed.region=count.x/count.y)

Armed2<-filter(Armed1,Armed != "Unknown", Armed != "Disputed")
ggplot(data=Armed2) +
  geom_col(aes(x=Armed,y=Prop.armed.region)) + 
  xlab("Armed") + ylab("Proportional of Deaths") + 
  theme_light()+
  facet_wrap(region~.~.)+
  theme(axis.text.x=element_text(angle=45,hjust=1))


## Adding gun law rankings

Armed.laws<-PoliceKillings %>%
  select(Armed,region) %>%
  left_join(gun.scorecard,by="region") %>%
  select(Armed,`2015 Grade`) %>%
  group_by(Armed,`2015 Grade`) %>%
  dplyr::summarize(count=n()) %>%
  ungroup() %>%
  left_join(laws.kills,by="2015 Grade") %>%
  mutate(armed.prop.by.laws=count/Kills.By.Laws) %>%
  group_by(Armed,`2015 Grade`)

# Figure of proportional deaths by armed status and gun laws
Armed.laws= filter(Armed.laws, Armed != "Unknown", Armed != "Disputed")
ggplot(data=Armed.laws) +
  geom_col(aes(x=Armed,y=armed.prop.by.laws)) + 
  xlab("Armed Status") + ylab("Proportion of Deaths") + 
  theme_light()+
  facet_wrap(Armed.laws$`2015 Grade`~.)+
  theme(axis.text.x=element_text(angle=45,hjust=1))
```

#Summary
PARAGRAPH 1:Throughout the analysis of the first 10 questions, we focused on attempting  to determine which of the factors in the initial dataset were most indicative of fatal confrontations by police officers. Those factors included: location, race, age, income, whether  the victim was armed, along with various census statistics for the area in which the victim was killed. Grouping the states by region in the U.S. elucidated the individual pieces of data on a larger scale, and allowed us to examine which geographic regions had the highest rates of police killings per 100,000 citizens. Furthermore, grouping enabled us to analyze the impact of race on police killings by region. In our analysis we also explored certain relationships between police killings and socioeconomic demographics such as  unemployment rates and the percentage of people in an area that attended college. There were a few unusual aspects of our data, such as how the observations for those who were shot in an airport had many missing values for the census information, leading us to remove these observations from the dataset before conducting analysis. Additionally, after investigating the cause of death variable in the dataset we found that a portion of the victims were actually not killed by firearms, but rather through means of taser and motor vehicle. An issue that arose during our analysis was that when we attempted to predict the likelihood of police killings, we realized this was not possible due to our data being limited to police confrontations that resulted in fatalities. In order to perform logistic regressions we would require aggregate data of all police confrontations, including those that did not result in death. 
	
The question we answered that left us with more questions was whether or not the victim of the killing was armed. This incited a follow up because different states have different carry laws in terms of how easy it is to have a gun.

PARAGRAPH 2: SHOULD SUMMARIZE WHAT YOU LEARNED FROM INVESTIGATING THE FOLLOW-UP QUESTIONS. WHY ARE THESE FOLLOW-UP QUESTIONS INTERESTING FOR INVESTIGATION? DESCRIBE THE TABLES/FIGURES YOU USED TO EXPLORE ANSWERS TO THESE FOLLOW-UP QUESTIONS? WHAT DID YOU LEARN FROM THE TABLES/FIGURES REGARDING THE FOLLOW-UP QUESTIONS YOU PROPOSED?














