---
title: "STOR320_Patchwork_test"
author: "Chris Jakuc"
date: "12/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(patchwork)
library(tidyverse)
library(dplyr)
library(Hmisc)
library(modelr)
library(rvest)
library(formattable)
```



```{r}
# Import Data Below
PoliceKillings<-read_csv("PKcleanP2.csv")

# Add DC to the South region
PoliceKillings$region[which(PoliceKillings$State=="DC")]<-"South"

# Import CSV
PoliceKillings.table<-read_csv("PoliceKillingsTable.csv")

```


```{r}
# Adding populations by state and adding populations by region
populations<-read_csv("StatePops.csv")
pops<-populations[-1,3:11]


# Import states by region
region.data<-read_csv("StatesbyCensusRegion.csv")


# Joining state pop and region data
pops1<-pops %>%
  select("GEO.display-label","respop72015") %>%
  rename(State="GEO.display-label",Pop2015=respop72015) %>%
  left_join(region.data,by="State") %>%
  mutate(Pop2015=as.numeric(Pop2015)) %>%
  group_by(region) %>%
  mutate(PopByRegion=sum(Pop2015)) %>%
  select(PopByRegion,region) %>%
  ungroup() %>%
  mutate(Pop.prop=PopByRegion/PopByRegion[1])

# Create a tibble of uniqe regions, pops, and prop pops
pops2<- tibble(PopByRegion=unique(pops1$PopByRegion),Region=unique(pops1$region),Pop.prop=unique(pops1$Pop.prop))

# Converting region variable to a factor
regions<-pops2 %>%
  mutate(Region=as_factor(Region))
levels(regions$Region)

# Remove the row for US total police killings
regions<-regions[-c(is.na(regions$Region)),]
```


```{r}
# Count how many police killings are in each state
state.kills<-PoliceKillings %>%
  select(State) %>%
  mutate(State=as_factor(State)) %>%
  group_by(State) %>%
  dplyr::summarize(count=n())

# Count how many police killings are in each region
region.kills<-PoliceKillings %>%
  select(region) %>%
  mutate(region=as_factor(region)) %>%
  group_by(region) %>%
  dplyr::summarize(count=n())
```


```{r}
# Create population state data frame
states<- pops %>%
  select("GEO.display-label","respop72015") %>%
  rename(State="GEO.display-label",Pop2015=respop72015) %>%
  mutate(Pop2015=as.numeric(Pop2015)) %>%
  left_join(region.data,by="State") %>%
  select(-region)

# Join police killings by state with population state data and create a variable for kills per 100,000 per state
states1<-states[-c(1),] %>%
  select(Pop2015,State,state) %>%
  right_join(state.kills,by=c("state"="State")) %>%
  mutate(kill.pc.state=count/Pop2015*100000)

# Join police killings by region with population regions data and create a variable for kills per 100,000 per region
regions1<-regions %>%
  select(PopByRegion,Region) %>%
  right_join(region.kills,by=c("Region"="region")) %>%
  mutate(kill.pc.region=count/PopByRegion*100000)
```


```{r}
Races <- PoliceKillings %>% 
  group_by(Race,region) %>%
  dplyr::summarize(count=n()) %>%
  ungroup() %>%
  mutate(prop=count/sum(count)) 

# Remove the row for unknown races
Races<-Races[-c(which(Races$Race=="Unknown")),]

NE= 47433806+38008094+6049443+3151094+6991969
MW= 52096633+6843367+388068+1738361+4661678
S= 68706462+21578475+728410+3238092+18227508
W= 38006363+3214563+1002967+6923296+20596439
total=NE+MW+S+W

# Proportions of each race that live in each region in 2015
# https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?src=bkmk

race.table<-data.frame("Region"= c("Northeast","Midwest","South","West"),
                           "White"=c(47433806,52096633,68706462,38006363),
                           "Black"=c(38008094,6843367,21578475,3214563),
                           "Native American"=c(6049443,388068,728410,1002967),
                           "Asian/Pacific Islander"=c((117653+3033441),(1713429+24932),(3170814+67278),(6547440+375856)),
                            "Hispanic"=c((55317240-48325271),(66927001-62265323),(114555744-96328236),(71945553-51349114)),
                           "Total"=c(NE,MW,S,W))

race.prop.table<-data.frame("Region"= c("Northeast","Midwest","South","West"),
                            "White"=c(47433806/NE,52096633/MW,68706462/S,38006363/W),
                            "Black"=c(38008094/NE,6843367/MW,21578475/S,3214563/W),
                            "Native American"=c(6049443/NE,388068/MW,728410/S,1002967/W),
                            "Asian/Pacific Islander"=c((117653+3033441)/NE,(1713429+24932)/MW,(3170814+67278)/S,(6547440+375856)/W),
                            "Hispanic"=c((55317240-48325271)/NE,(66927001-62265323)/MW,(114555744-96328236)/S,(71945553-51349114)/W),
                            "Total"=c(NE/total,MW/total,S/total,W/total))

race.rank<-race.prop.table %>%
  rename("Native American"=Native.American,
         "Asian/Pacific Islander"=Asian.Pacific.Islander,
         "Hispanic/Latino"=Hispanic) %>%
  gather(White:`Hispanic/Latino`,key="Race",value="Race_Proportion_by_Region",factor_key = T) %>%
  mutate(Region=factor(Region)) %>%
  filter(!(Race =="Native American"&Region!="West"))

# Ranks of the proportion of race for all the regions
ranks<-c(4,2,3,1,3,1,2,4,3,4,2,1,1,1,4,2,3)

# Plotting proportion of police killings in each region by race
Races1<-Races %>%
  left_join(region.kills,by="region") %>%
  mutate(Prop.race.region=count.x/count.y) %>%
  left_join(race.rank,by=c("region"="Region","Race")) %>%
  cbind(ranks)
```


```{r}
# Mean Median Income for entire regions (including living people) estimated by census bureau as of 2015 from https://www.census.gov/library/publications/2016/demo/p60-256.html

region.names<-c("Northeast","Midwest","South","West")
region.mean.inc<-c(62182,57082,51174,61442)


mean.inc<-tibble(region=region.names,mean=region.mean.inc)
```


```{r}
#count of whether each killing occured in an area that has an unemployment rate higher than the national average of 3.6%
PK1=PoliceKillings%>%
  mutate(higherthanavg=if_else(Unemployment_Rate>0.054,"Yes","No"))
table(PK1$higherthanavg) 

#splitting up unemployment into ranges and then finding the count for each bin
URranges1=cut2(PoliceKillings$Unemployment_Rate,c(0.02,0.04,0.06,0.08,0.1,0.12,0.14,0.16,0.18,0.2,0.22,0.24,0.26,0.28,0.3,0.32,0.34,0.36,0.38,0.4)) 
summary(URranges1) 
```


```{r}
# Web scrape the gun laws table from
URL.gun.scorecard = "https://lawcenter.giffords.org/scorecard2015/"
gun.scorecard = URL.gun.scorecard %>%
                      read_html() %>%
                      html_table(fill=1) %>%
                      .[[1]]

# Removing the minus and plus signs from the grades
gun.scorecard$`2015 Grade`[gun.scorecard$`2015 Grade` %in% "A-"]="A"
gun.scorecard$`2015 Grade`[gun.scorecard$`2015 Grade` %in% "B+"]="B"
gun.scorecard$`2015 Grade`[gun.scorecard$`2015 Grade` %in% "B-"]="B"
gun.scorecard$`2015 Grade`[gun.scorecard$`2015 Grade` %in% "C-"]="C"
gun.scorecard$`2015 Grade`[gun.scorecard$`2015 Grade` %in% "D-"]="D"

# Adding the region
gun.scorecard<-gun.scorecard %>%
  left_join(region.data,by="State")
 

gun.scorecard<-pops %>%
  select("GEO.display-label","respop72015") %>%
  rename(State="GEO.display-label",Pop2015=respop72015) %>%
  left_join(gun.scorecard,by="State") %>% 
  mutate(Pop2015=as.numeric(Pop2015)) %>%
  left_join(state.kills,by=c("state"="State"))

# Remove DC and Total US
gun.scorecard1<-gun.scorecard[-c(which(is.na(gun.scorecard$`2015 Grade`))),] %>%
  select(-`Death Rate Per 100k`,-`New Smart Gun Laws in 2015`)

# Turning the gun law grades into factors to run a regression later
gun.scorecard2<-gun.scorecard1[-c(which(is.na(gun.scorecard1$count))),]

gun.scorecard3<-gun.scorecard2 %>%
  mutate("2015 Grade"=factor(gun.scorecard2$`2015 Grade`)) %>%
  mutate(KillCount=count) %>%
  select(-count)

# Find how many unique groups of gun laws there are and how many kills per capita are in each
gun.scorecard4<-gun.scorecard3 %>%
  group_by(`2015 Grade`) %>%
  dplyr::summarize(count.gunlaws.ranks=n(),
                   KillCount=sum(KillCount),
                   Population=sum(Pop2015),
                   "Kill.per.100,000"=KillCount/Population*100000) %>%
  ungroup()
  
  
gun.table<-gun.scorecard4 %>%
  rename("Number of States"=count.gunlaws.ranks,"Kill Count"=KillCount,"Kill Rate per 100,000 People"=`Kill.per.100,000`)
```


```{r}
library(mapproj)
library(usmap)
library(mapdata)
library(sf)

PoliceKillingsMap = 
  PoliceKillings %>% 
  filter(State != "HI") %>%
  filter(State != "AK")

labs1 <- data.frame(
long = PoliceKillingsMap$Longitude, 
lat = PoliceKillingsMap$Latitude, 
state = PoliceKillingsMap$State,
stringsAsFactors = FALSE)


us_states1 = map_data("state")

us_states2 = 
  us_states1 %>% 
  mutate(State = region) %>% 
  select(-subregion,-region)

gun.scorecard[[1]] = tolower(gun.scorecard[[1]])

us_states3 = right_join(us_states2,gun.scorecard)

rename(us_states3, Rank = "2015 Grade") 


labs2 = right_join(us_states3, labs1)


one = ggplot(data = us_states3, aes(x =long, y = lat, group = group, fill = Rank)) + guides() + coord_fixed(1.3)

two = one + geom_polygon(color = "gray90", size = 0.1) + 
  coord_map(projection = "albers", lat0 = 39, lat1 = 45)

three = two + scale_fill_gradient(low = "white", high = "skyblue1")


four = three +
  geom_point(data = labs2, aes(x = long, y = lat),color="red", alpha = 0.4, size = 0.5) + 
  # ggtitle("Police Killings in States with Ranked Gun Laws")+
  theme_dark()+
  theme(plot.background = element_rect(fill = "lightgray"))+
  xlab("Longitude")+
  ylab("Latitude")
```

**Figure One: Proportion of Deaths per Region by Median Household Income**
```{r, echo=FALSE, warning=FALSE}
p1 <- ggplot(PoliceKillings)+
  geom_histogram(aes(x=med.tract.household.inc),bins=20,fill='skyblue1')+
  facet_wrap(region~.)+
  geom_vline(aes(xintercept=mean),data=mean.inc,linetype="dashed",color="white")+
  theme_dark()+
  theme(plot.background = element_rect(fill = "lightgray"))+
  xlab("Median Household Income")+
  ylab("Count of Deaths")+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  scale_x_continuous(breaks=c(25000,50000,75000,100000))+
  # ggtitle("Proportion of Deaths per Region by Median Household Income")+
  labs(subtitle = "Dashed Line is Regional Mean Median Household Income")
```

**Figure Two: Regional Proportion of Police Killings by Race**
```{r, echo=FALSE, warning=FALSE}
p2 <- ggplot(Races1)+
  geom_col(aes(x=Race,y=Prop.race.region),fill="skyblue1")+
  geom_text(aes(x=Race,y=Prop.race.region,label=ranks),color="white")+
  facet_wrap(region~.)+
  ylab("Proportion of police killings by region")+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  # ggtitle("Regional Proportion of Police Killings by Race")+
  theme_dark()+
  theme(plot.background = element_rect(fill = "lightgray"))+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  labs(subtitle="Ranked by Total Race Proportion in All Regions")
```

**Table Two: Proportions of Populatiuon by Race and Region**
```{r, echo=FALSE, warning=FALSE}
p3 <- formattable(race.prop.table)

```

**Table Three: Measures of Error within each Model**
```{r, echo=FALSE, warning=FALSE}
p4 <- formattable(table2)
```

**Figure Five: Police Killings per 100,000 People by Grade of Gun Laws**
```{r, echo=FALSE, warning=FALSE}
p5 <- ggplot(gun.scorecard4)+
  geom_col(aes(x=`2015 Grade`,y=`Kill.per.100,000`),fill="skyblue1")+
  xlab("Gun Control Law Rankings")+
  ylab("Police Killings per 100,000 People")+
  # ggtitle("Police Killings per 100,000 People by Grade of Gun Laws")+
  theme_dark()+
  theme(plot.background = element_rect(fill = "lightgray"))
```


**Table Five: Police Killing Rates by Gun Law Grades**
```{r, echo=FALSE, warning=FALSE}
p6 <- formattable(gun.table)
```

**Figure Six: Police Killings in States with Ranked Gun Laws**
```{r, echo=FALSE, warning=FALSE}
p7 <- four
```


```{r}
(p5) /
  p7
```


```{r}
(p1 | p2)
```

